---
title: "Shiny Calendar Heatmap Demo"
author: "SportSciData"
date: "15/04/2019"
output: html_document
runtime: shiny
css: mystyle.css
---
Don't forget to check Parts [I](https://www.sportscidata.com/2019/04/04/how-to-create-interactive-reports-with-r-markdown-part-i/), [II](https://www.sportscidata.com/2019/04/12/using-data-visualisation-in-r-markdown/) and [III](https://www.sportscidata.com/2019/04/19/how-to-create-interactive-reports-in-r-markdown-shiny-reactive-elements/) of the R Markdown series where we covered producing *tables of data, data visualisation and most recently, how to add reactive elements to your report!*

Below you will see a reactive calendar heatmap which produces a table underneath based on on the date selected from the calendar. Different metrics, players and date ranges are availalble to choose from. 

***

# Calendar Heatmap & Reactive Table
The calendar was produced using [ggiraph](https://davidgohel.github.io/ggiraph/articles/offcran/using_ggiraph.html) and is based off post from [John Mackintosh](https://www.johnmackintosh.com/2016-11-27-calendar-heatmaps/). The table is a combination of [Kable](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html) and [formatTable](https://renkun-ken.github.io/formattable/). 


```{r setup, include=FALSE}
library(tidyverse)
library(htmlwidgets)
library(plotly)
library(ggiraph)
library(gridExtra)
library(RColorBrewer)
library(dygraphs)
library(xts)
library(lubridate)
library(shiny)
library(formattable)
library(kableExtra)
library(magrittr)
library(shinycssloaders)
library(shinyjs)
library(shinyalert)

gps_data <- readr::read_csv('trial_data.csv')

df2 <- gps_data %>%
  select(Date, Name, dist_total, msr_total, hsr_total, percentVmax) %>%
  gather(metric, value, -Name, -Date) %>%
  mutate(
    metric = factor(metric, levels = c('dist_total', 'msr_total', 'hsr_total',
                                       'percentVmax'),
                    labels = c('Total Distance', 'Moderate Speed Distance', 'High Speed Distance',
                               'Percentage of Max Velocity'))
  )


tooltips_func <- function(metric){
  
  if_else(metric == 'Percentage of Max Velocity',
         paste0(df2$Date, "\n", 'Percent Max Velocity ', round(df2$value*100, 2), "%"),
         paste0(df2$Date, "\n", 'Distance ', round(df2$value , 0), 'm')
  )
}

ts_data2 <- df2 %>%
  mutate(
    Date = anytime::anydate(Date),
    dow = wday(Date,label=TRUE, week_start = 1),
    dow = factor(dow, levels = rev(levels(dow))),
    week = week(Date),
    weeks = format(Date, "%W"),
    weeks = factor(weeks, levels = unique(weeks)),
    weekStart = Date - as.POSIXlt(Date)$wday,
    month = month(Date, label = TRUE, abbr = FALSE),
    month = factor(month, levels = c(
      'January', 
      'February', 
      'March', 
      'April', 
      'May', 
      'June', 
      'July', 
      'August', 
      'September', 
      'October', 
      'November', 
      'December'
    )),
    year = year(Date),
    year = factor(year),
    yearmonth = paste(month, year),
    tooltips = tooltips_func(metric)
  ) %>% arrange(year, month) %>%
  mutate(yearmonth = factor(yearmonth, levels = unique(yearmonth)))


tooltips_colour_func2 <- function(df, ...){
  
  td_scale <-  colorRampPalette(c('#08B3E5', '#2AF598', '#00FF3E', 'orange', 'red')) (nlevels(
    factor(df$value)))
  
  vmax_scale <- colorRampPalette(c('red', 'orange', 'lightgreen', 'darkgreen')) (nlevels(
    factor(df$value)))
  
  if_else(df$metric == 'Percentage of Max Velocity', vmax_scale, td_scale)
}  

title_func <- function(df){ ###Sorted
  
  title_td <- expression(paste(bold('Total Distance '), ' Covered Per Day'))
  title_msr <- expression(paste(bold('Moderate Speed Distance '), ' Covered Per Day'))
  title_hsr <- expression(paste(bold('High Speed Distance '), ' Covered Per Day'))
  title_vmax <- expression(paste(bold('Percent Max Velocity '), ' Achieved Per Day'))
  
  if_else(df$metric == 'Total Distance',
         title_td,
         if_else(df$metric == 'Moderate Speed Distance',
                title_msr,
                if_else(df$metric == 'High Speed Distance',
                       title_hsr,
                       title_vmax)))
}


subtitle_func <- function(df){
  
  sub_dist <-  paste0('Blue ->> Low, Green ->> OK, Orange ->> Warning, Red ->> High\n','Athlete Name: ', df$Name)
  sub_vel <-  paste0('Red ->> Low, Orange ->> OK, Green ->> Good\n' , 'Athlete Name: ', df$Name)
  
  if_else(df$metric == 'Percentage of Max Velocity',
         sub_vel,
         sub_dist)
}


appCSS <- "
#loading-content {
  position: absolute;
  background: #000000;
  opacity: 0.9;
  z-index: 100;
  left: 0;
  right: 0;
  height: 100%;
  text-align: center;
  color: #FFFFFF;
}
"

```


```{r, echo=FALSE}


shinyApp(
  
  ui = fluidPage(
    
    useShinyjs(),
    useShinyalert(),
    inlineCSS(appCSS),
    
    div(
      id = "loading-content",
      h2("Loading...")
    ),
    
    dateRangeInput('date_select2', label = 'Select Date Range', 
                   start = min(ts_data2$Date),
                   end = max(ts_data2$Date),
                   min = min(ts_data2$Date),
                   max = max(ts_data2$Date),
                   weekstart = 1
    ),
    
    selectInput('name_select', 
                label = 'Select Name', 
                choices = unique(df2$Name)
    ),
    
    selectInput('metric_select', 
                label = 'Select Metric Name', 
                choices = unique(df2$metric),
                selected = 'Total Distance'
    ),
    
    ggiraphOutput("calHeat") %>% withSpinner(color="#0dc5c1"),
    downloadButton('downloadFile', 'Download', class='dlbutton'),
    shiny::htmlOutput('dt') %>% withSpinner(color="#0dc5c1")
    
    
  ),
  
  server = function(input, output, session){
    
    shinyalert(
      title = "Hello",
      text = "This is to show annoying popups are possible here!",
      type = "success",
      confirmButtonText = "OK",
    )
    
    
    output$calHeat = renderggiraph({
      
      ts_data3 <- ts_data2 %>%
        filter(Date >= input$date_select2[1] & Date <= input$date_select2[2]  & 
                 Name == input$name_select & metric == input$metric_select)
      
      x2 <- ts_data3 %>%
        ggplot(aes(weekStart, dow, fill = factor(value))) +
        geom_tile_interactive(colour = "white", 
                              aes(tooltip = tooltips,
                                  data_id = as.character(Date)), size = .1) + 
        scale_fill_manual(values = 
                            tooltips_colour_func2(ts_data3)) +
        scale_x_date(date_breaks = "1 week",date_labels="%d-%b") +
        bbplot::bbc_style()+
        ggExtra::removeGrid() + 
        ggtitle(title_func(ts_data3),
                subtitle = subtitle_func(ts_data3)) +
        labs(x = "Week Beginning", y = NULL) +
        facet_wrap(~ yearmonth, scales = 'free_x') +
        theme(
          plot.title = element_text(hjust = 0, size = 14), 
          plot.subtitle = element_text(size = 10),
          strip.text = element_text(size = 6),
          axis.ticks = element_blank(), 
          axis.text.y = element_text(size = 4), 
          axis.text.x = element_text(size = 5, angle = 70), 
          legend.position = "none")
      
      x2 <- girafe(code = print(x2))
      
      tooltip_css <- "padding:10px;border-radius:10px 20px 10px 20px;"
      
      girafe_options(x2, 
                     opts_tooltip(offx = -60, 
                                  offy = -120, use_fill = TRUE,
                                  css = tooltip_css),
                     opts_selection(css = 'fill:grey'))
    })
    
    output$dt <- shiny::renderUI({
      
      if(is.null(input$calHeat_selected)){
        print('Select Dates to View Table')
      } else{

      out <- ts_data2 %>%
        filter(Name == input$name_select &
                 Date %in% anytime::anydate(input$calHeat_selected)) %>%
        select(1:4) %>% 
        spread(key = metric, value = value) %>%
        mutate_at(vars(3:5), round) %>%
        mutate(`Percentage of Max Velocity` = 
                 round(`Percentage of Max Velocity`, 2)) %>%
        mutate(
          `High Speed Distance` = ifelse(
            `High Speed Distance` > 150,
            cell_spec(`High Speed Distance`, "html", color = "red", bold = T),
            ifelse(`High Speed Distance` < 50,
                   cell_spec(`High Speed Distance`,"html",
                             color = "blue", bold = T),
                   cell_spec(`High Speed Distance`, "html",
                             color = "green", italic = T))),
          `Percentage of Max Velocity` =
            cell_spec(
              `Percentage of Max Velocity`, bold = T,
              color = spec_color(`Percentage of Max Velocity`, end = 0.9),
              font_size = spec_font_size(`Percentage of Max Velocity`)),
          `Total Distance` = color_bar("#FA614B")(`Total Distance`),
          `Moderate Speed Distance` =
            color_tile("#DeF7E9", "#71CA97")(`Moderate Speed Distance`)
        ) %>% kable('html',  caption = 'Session Loads', escape=FALSE,
                    align = c(rep('l', 3), rep('c', 3))) %>%
        kable_styling( bootstrap_options = c('striped', 'hover', 'responsive', 'condensed'))
      
      HTML(out)      }
    })  
    
    
    df <- reactive({
      ts_data2 %>%
        filter(Name == input$name_select &
                 Date %in% anytime::anydate(input$calHeat_selected)) %>%
        select(1:4) %>% 
        spread(key = metric, value = value) %>%
        mutate_at(vars(3:5), round) %>%
        mutate(`Percentage of Max Velocity` = 
                 scales::percent(`Percentage of Max Velocity`))
      
      
    })
    
    
    output$downloadFile <- downloadHandler(
      filename = function(){paste0(input$name_select,"_", Sys.Date(), ".csv")},
      content = function(file) {
        write.csv(df(), file,
                  row.names = FALSE)
      }
    )
    
    # Hide the loading message when the rest of the server function has executed
    hide(id = "loading-content", anim = TRUE, animType = "fade")    
    show("app-content")
  },
  
  options = list(height = 1000)
)
```