---
title: "Shiny Report Demo"
author: "SportSciData"
date: "15/04/2019"
output: 
  html_document:
    toc: true
runtime: shiny
css: reportcss.css
---

Don't forget to check Parts [I](https://www.sportscidata.com/2019/04/04/how-to-create-interactive-reports-with-r-markdown-part-i/) and [II](https://www.sportscidata.com/2019/04/12/using-data-visualisation-in-r-markdown/) of the R Markdown series where we covered producing *tables of data* first and then *data visualisation*


```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(ggiraph)
library(dygraphs)
library(lubridate)
library(shiny)
library(formattable)
library(kableExtra)
library(magrittr)
library(DT)
library(sparkline)
library(colormap)


knitr::opts_chunk$set(echo = TRUE)
gps_data <- readr::read_csv('trial_data.csv')
hdsts <- list.files(path="headshots", pattern = ".png", full.names = TRUE,
                    all.files = TRUE)

spk_tool <- function(labels) {
  htmlwidgets::JS(
    sprintf(
      "function(sparkline, options, field){
  return %s[field[0].offset];
}",
jsonlite::toJSON(labels)
    )
  )
}


```

-------------------------

# Name & Date Range Selection

```{r cars, echo=FALSE}

selectInput('name_select', 
            label = 'Select Name', 
            choices = unique(gps_data$Name))

dateRangeInput('date_select2', 
               label = 'Select Date Range', 
               start = max(gps_data$Date)-days(60),
               end = max(gps_data$Date)-days(30),
               min = min(gps_data$Date),
               max = max(gps_data$Date),
               weekstart = 1)

```

## [Kable](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html) based table produced from selections above

```{r, echo=FALSE}

renderUI({
  
  gps_data %>%
    select(Date, Name, dist_total, msr_total, hsr_total, percentVmax) %>%
    filter(Date >= input$date_select2[1] & Date <= input$date_select2[2]  & 
             Name == input$name_select) %>%
    arrange(desc(Date)) %>%
    mutate(dist_total =  round(dist_total, 0), msr_total = round(msr_total, 0),
           hsr_total = round(hsr_total, 0),percentVmax = round(percentVmax, 2)) %>%
    mutate(dist_total = cell_spec(dist_total, "html", 
                                  color = ifelse(dist_total > 4000, "red", "blue")),
           msr_total = ifelse(msr_total > 1000,
                              cell_spec(msr_total, "html", color = "red", bold = T),
                              ifelse(msr_total < 400,
                                     cell_spec(msr_total,"html", color = "blue", bold = T),
                                     cell_spec(msr_total, "html", color = "green", italic = T))),
           percentVmax = cell_spec(percentVmax, bold = T, color = spec_color(percentVmax, end = 0.9),
                                   font_size = spec_font_size(percentVmax))) %>%
    kable(escape=FALSE, caption = 'Session Loads', 
          align = c('l','l', 'l', 'l', 'l', 'c', 'c', 'c','c')) %>%
    kable_styling( bootstrap_options = c('striped', 'hover', 'responsive', 
                                         'condensed')) %>%
    column_spec(1, width = '5cm', border_right = TRUE, bold = TRUE, background = "grey") %>%
    HTML()
  
  
})
```

---

## [Plotly](https://plot.ly/r/) plot based on name selected above

```{r, echo=FALSE}

renderPlotly({
  
  plot_title <- input$name_select
  
  p <- gps_data %>%
    select(Date, Name, dist_total, msr_total, hsr_total) %>%
    filter(Date >= input$date_select2[1] & Date <= input$date_select2[2]  & 
             Name == input$name_select) %>%
    arrange(desc(Date)) %>%
    mutate(
      Distance = round(dist_total, 0),
      Date = factor(Date)
    ) %>% gather(metric, value, -Date, -Name) %>%
    mutate(
      metric=factor(metric, levels = c('dist_total', 'msr_total', 'hsr_total'),
                    labels = c('Total Distance', 'Moderate Speed Distance',
                               'High Speed Distance'))
    ) %>% filter(complete.cases(.)) %>%
    ggplot(aes(Date, value, colour = value, shape=metric, 
               text=paste0(Date, '\n', metric, ": ", round(value,0), 'm')))+
    geom_point() +
    labs(title = plot_title)+
    labs(x ='Date', y ='Distance (m)') +
    scale_color_gradient2(low = 'blue', mid = 'green', high = 'red') +
    ylim(0,10000)+
    bbplot::bbc_style()+
    theme(
      legend.text = ggplot2::element_text(size = 10),
      axis.text.x = element_text(size = 12, angle = 70, colour = 'azure4'),
      axis.text.y = element_text(size = 12, colour = 'azure4')
    )
  
  ggplotly(p, tooltip = 'text') %>% layout(hovermode = 'compare')
  
})
```

---

# Date Selection

```{r cars2, echo=FALSE}

df_date <- enframe(seq.Date(min(gps_data$Date), max(gps_data$Date), by=1)) %>%
  anti_join(gps_data, by=c('value'='Date'))

dateInput('date_select', 
          label = 'Select Date', 
          value = "2018-05-01",
          min = min(gps_data$Date),
          max = max(gps_data$Date),
          weekstart = 1,
          datesdisabled = df_date$value)


```

## [Formattable](https://renkun-ken.github.io/formattable/) and [Kable](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html) based table based off date selection

```{r, echo=FALSE}

renderUI({
  
  data2 <-  gps_data %>%
    select(Date,week,`Forward/Back`, `Position Specific`, Name, dist_total, msr_total, hsr_total, percentVmax) %>%
    filter(Date == input$date_select) %>%
    mutate(`Forward/Back` = factor(`Forward/Back`, levels = c('Forwards', 'Backs')),
           `Position Specific` = factor(`Position Specific`, levels = c(
             'Tighthead Prop', 'Hooker', 'Loosehead Prop', 'Second Row', 'Back Row',
             'Scrum Half','Fly Half', 'Centre', 'Wing', 'Full Back')
           )) %>% arrange(`Forward/Back`, `Position Specific`)
  
  pos_groups <- table(data2$`Forward/Back`)
  
  data2 %>%
    select(-`Forward/Back`) %>%
    mutate(
      dist_total =  round(dist_total, 0), 
      msr_total = round(msr_total, 0),
      hsr_total = round(hsr_total, 0),
      percentVmax = round(percentVmax, 2),
      hsr_total = ifelse(hsr_total > 150,
                         cell_spec(hsr_total, "html", color = "red", bold = T),
                         ifelse(hsr_total < 50,
                                cell_spec(hsr_total,"html", color = "blue", bold = T),
                                cell_spec(hsr_total, "html", color = "green", italic = T))),
      percentVmax = cell_spec(percentVmax, bold = T, color = spec_color(percentVmax, end = 0.9),
                              font_size = spec_font_size(percentVmax)),
      `dist_total` = color_bar("#FA614B")(dist_total),
      msr_total = color_tile("#DeF7E9", "#71CA97")(msr_total)) %>%
    kable('html',  caption = 'Session Loads', escape=FALSE, 
          align = c(rep('l', 5), rep('c', 4)), 
          col.names = c('Date', 'Week','Position', 'Name','Total Distance',
                        'Mod Speed Dist', 'High Speed Dist', 'Percent Max Vel')) %>%
    kable_styling( bootstrap_options = c('striped', 'hover', 'responsive', 'condensed')) %>%
    pack_rows(index = setNames(pos_groups, names(pos_groups)),
              label_row_css = "background-color: #666; color: #fff;") %>%
    HTML()
  
  
})


```

## [ggiraph](https://github.com/cardiomoon/ggiraphExtra/tree/master/R) based off date selection 

```{r, echo=FALSE}

renderggiraph({
  
  td_plot <- gps_data %>% 
    filter(Date ==  input$date_select) %>% 
    mutate(Name = forcats::fct_reorder( Name, dist_total, desc=T),
      tooltip = paste0(Name, "\n",`Position Specific`, "\n", round(dist_total, 0),"m")) %>%
    ggplot(aes(Name, dist_total, fill = `Position Specific`, 
               label = paste0(round(dist_total, 0),"m"))) +
    ggiraph::geom_bar_interactive(aes(y=dist_total, fill=`Position Specific`, 
                                      tooltip=tooltip),stat = 'identity') +
    coord_flip() + geom_text(hjust=1, size=2) + bbplot::bbc_style() +
    labs(y = "Total Distance", x = 'Player Name', 
         fill='Position', title = 'Total Distance Covered', type='Garamond') +
    theme(axis.text.x = element_blank(),axis.text.y = element_text(size=12),
      axis.ticks.x = element_blank(),
      legend.text = element_text(size=4.5, color="#222222"),
      legend.justification=c(1,0), 
      legend.position=c(1,0),
      legend.spacing.x = unit(0.06, 'cm'),
      plot.title = element_text(size=10))
  hsr_plot <- gps_data %>% 
    filter(Date ==  input$date_select) %>% 
    mutate(Name = forcats::fct_reorder( Name, dist_total, desc=T),
      tooltip = paste0(Name, "\n",`Position Specific`, "\n", round(hsr_total, 0),"m")) %>%
    ggplot(aes(Name, hsr_total, , label= paste0(round(hsr_total, 0),"m"))) +
    ggiraph::geom_bar_interactive(aes(y=hsr_total, fill=`Position Specific`,
                                      tooltip=tooltip),stat = 'identity') +
    coord_flip() +geom_text(hjust=1, size=2) +bbplot::bbc_style() +
    labs(y = "High Speed Distance", x = 'Player Name', 
         fill='Position', title = 'High Speed Distance Covered', type='Garamond') +
    theme( axis.text.x = element_blank(),axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      legend.position = "none",
      plot.title = element_text(size=10))
  ggiraph_plots <- cowplot::plot_grid(td_plot, hsr_plot)
  
  plots2 <- ggiraph::girafe(code=print(ggiraph_plots))
  tooltip_css <- "padding:10px;border-radius:10px 20px 10px 20px;"
  girafe_options(plots2, opts_tooltip(offx = -60, offy = -120, use_fill = TRUE,
                              css=tooltip_css),opts_zoom(max = 5))
})

```

---


# Date slider input

```{r echo=FALSE}

sliderInput('date_select3', 
            label = 'Select Date Range', 
            max  = max(gps_data$Date),
            min = min(gps_data$Date),
            timeFormat = '%F',
            value = c( max(gps_data$Date)-days(60), max(gps_data$Date)),
            step = days(1))

selectInput('pos_select', 
            label = 'Select Position', 
            choices = unique(gps_data$`Position Specific`),
            multiple = TRUE,
            selected = 'Hooker')


```

## [DT](https://rstudio.github.io/DT/) plus [sparklines](https://github.com/htmlwidgets/sparkline) colour coded by value based off date slider and positional input *with CSS to design tooltip*

```{r, echo=FALSE}
renderDT({
  gps_data %>%
    filter(Date >= input$date_select3[1] & Date <= input$date_select3[2] &
             `Position Specific` %in% input$pos_select) %>%
    group_by(Name) %>% 
    arrange(Date) %>%
    summarize(
      Distance = spk_chr(
        round(dist_total,0),
        type='bar',
        width = 120,
        height = 35),
      MSR = spk_chr(
        round(msr_total,0),
        type='bar',
        width = 160,
        height = 40),
      HSR = spk_chr(
        round(hsr_total,0),
        type='bar',
        width = 160,
        height = 40),
      Percent_Velocity = sparkline::spk_chr(
        round(percentVmax,2),
        type='bar',
        width = 160,
        height = 40)
    ) %>%
    DT::datatable(escape = F, rownames = F,  
                  options = list(
                    pageLength = n_distinct(gps_data$Name), 
                    fnDrawCallback = 
                      htmlwidgets::
                      JS('function(){HTMLWidgets.staticRender();}'))) %>% 
    spk_add_deps()
  
})

```


